---
title: "Jeff"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# Cleaning

```{r}
data <- read.csv(here("data", "data_export_20240424.csv"), sep = ";")
```

```{r}
consumptiondata <- read.csv(here("data", "energyconsumption.csv"), sep = ";")
```

```{r}
temperature <- read.csv(here("data", "temperature.csv"), sep = ";")
```

## Convert Month into Date for the data dataset

```{r}
data$Month <- as.Date(data$Month, format = "%d.%m.%y")
```

## Convert Month into Date for the consumptiondata dataset

```{r}
consumptiondata$Month <- as.Date(consumptiondata$Month, format = "%d.%m.%y")
```

## Pivot Months for the temperature dataset

```{r}
temperature <- temperature %>% 
  pivot_longer(c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec'),
  names_to = "Month",
  values_to = "Temperature")
```

## Replace "Not Available" with 0 (Production)

```{r}
data$Solar_Energy_Production <- replace(data$Solar_Energy_Production, data$Solar_Energy_Production == "Not Available", 0)
data$Solar_Energy_Production %>% as.numeric(data$Solar_Energy_Production)
typeof(data$Solar_Energy_Production)

data$Solar_Energy_Production <- ifelse(data$Solar_Energy_Production == "Not Available", 0, data$Solar_Energy_Production)
data$Wind_Energy_Production <- ifelse(data$Wind_Energy_Production == "Not Available", 0, data$Wind_Energy_Production)

data$Solar_Energy_Production <- as.numeric(data$Solar_Energy_Production)
data$Wind_Energy_Production <- as.numeric(data$Wind_Energy_Production)
```

## Replace "Not Available" into 0 (Consumption)
```{r}

consumptiondata$Solar_Energy_Consumption <- replace(consumptiondata$Solar_Energy_Consumption, consumptiondata$Solar_Energy_Consumption == "Not Available", 0)
consumptiondata$Solar_Energy_Consumption <- as.numeric(consumptiondata$Solar_Energy_Consumption)
typeof(consumptiondata$Solar_Energy_Consumption)


consumptiondata$Solar_Energy_Consumption <- ifelse(consumptiondata$Solar_Energy_Consumption == "Not Available", 0, consumptiondata$Solar_Energy_Consumption)
consumptiondata$Wind_Energy_Consumption <- ifelse(consumptiondata$Wind_Energy_Consumption == "Not Available", 0, consumptiondata$Wind_Energy_Consumption)


consumptiondata$Solar_Energy_Consumption <- as.numeric(consumptiondata$Solar_Energy_Consumption)
consumptiondata$Wind_Energy_Consumption <- as.numeric(consumptiondata$Wind_Energy_Consumption)
```

## Switch from character to numeric (Consumption)
```{r}
consumptiondata <- consumptiondata %>%
  mutate(across(where(is.character), ~ as.numeric(gsub(",", ".", ., fixed = TRUE)))) %>%  
  filter(year(Month) != 2024)
```

## Switch from character to numeric (Temperature)

```{r}
month_map <- setNames(1:12, c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
temperature <- temperature %>%
  mutate(
    Month = month_map[Month],
    Temperature = as.numeric(gsub(",", ".", Temperature)) 
  )

sum(is.na(temperature$Month))

temperature <- temperature %>%
  mutate(Date = make_date(Year, Month, 1)) %>%  # Create Date assuming the 1st day of the month
  filter(Year >= 1973, Year != 2024) %>% 
  select(-Year, -Month)
```

# Plots

```{r}
## Coal

coal <- ggplot() + 
  # Plot Coal Production
  geom_line(data = data, aes(x = Month, y = Coal_Production, color = "Production")) + 
  # Plot Coal Consumption
  geom_line(data = consumptiondata, aes(x = Month, y = Coal_Consumption, color = "Consumption")) + 
  # Customize the plot
  labs(title = "Coal Production and Consumption", x = "Date", y = "Value") + 
  scale_color_manual(values = c("Production" = "blue", "Consumption" = "red"), name = "Type") +
  theme_minimal()

## Total Fossil Fuels Production

fossil_fuels <- ggplot() + 
  # Plot Total Fossil Fuels Production
  geom_line(data = data, aes(x = Month, y = Total_Fossil_Fuels_Production, color = "Production")) + 
  # Plot Total Fossil Fuels Consumption
  geom_line(data = consumptiondata, aes(x = Month, y = Total_Fossil_Fuels_Consumption, color = "Consumption")) + 
  # Customize the plot
  labs(title = "Total Fossil Fuels Production and Consumption", x = "Date", y = "Value") + 
  scale_color_manual(values = c("Production" = "blue", "Consumption" = "red"), name = "Type") +
  theme_minimal()

## Nuclear Electric Power

nuclear_electric <- ggplot() + 
  # Plot Nuclear Electric Power Production
  geom_line(data = data, aes(x = Month, y = Nuclear_Electric_Power_Production, color = "Production")) + 
  # Plot Nuclear Electric Power Consumption
  geom_line(data = consumptiondata, aes(x = Month, y = Nuclear_Electric_Power_Consumption, color = "Consumption")) + 
  # Customize the plot
  labs(title = "Nuclear Electric Power Production and Consumption", x = "Date", y = "Value") + 
  scale_color_manual(values = c("Production" = "blue", "Consumption" = "red"), name = "Type") +
  theme_minimal()

## Hydroelectric Power

hydroelectric <- ggplot() + 
  # Plot Hydroelectric Power Production
  geom_line(data = data, aes(x = Month, y = Hydroelectric_Power_Production, color = "Production")) + 
  # Plot Hydroelectric Power Consumption
  geom_line(data = consumptiondata, aes(x = Month, y = Hydroelectric_Power_Consumption, color = "Consumption")) + 
  # Customize the plot
  labs(title = "Hydroelectric Power Production and Consumption", x = "Date", y = "Value") + 
  scale_color_manual(values = c("Production" = "blue", "Consumption" = "red"), name = "Type") +
  theme_minimal()

## Solar Energy

solar <- ggplot() + 
  # Plot Solar Energy Production
  geom_line(data = data, aes(x = Month, y = Solar_Energy_Production, color = "Production")) + 
  # Plot Solar Energy Consumption
  geom_line(data = consumptiondata, aes(x = Month, y = Solar_Energy_Consumption, color = "Consumption")) + 
  # Customize the plot
  labs(title = "Solar Energy Production and Consumption", x = "Date", y = "Value") + 
  scale_color_manual(values = c("Production" = "blue", "Consumption" = "red"), name = "Type") +
  theme_minimal()

## Biomass Energy

biomass <- ggplot() + 
  # Plot Biomass Energy Production
  geom_line(data = data, aes(x = Month, y = Biomass_Energy_Production, color = "Production")) + 
  # Plot Biomass Energy Consumption
  geom_line(data = consumptiondata, aes(x = Month, y = Biomass_Energy_Consumption, color = "Consumption")) + 
  # Customize the plot
  labs(title = "Biomass Energy Production and Consumption", x = "Date", y = "Value") + 
  scale_color_manual(values = c("Production" = "blue", "Consumption" = "red"), name = "Type") +
  theme_minimal()

## Total Renewable Energy

renewable <- ggplot() + 
  # Plot Total Renewable Energy Production
  geom_line(data = data, aes(x = Month, y = Total_Renewable_Energy_Production, color = "Production")) + 
  # Plot Total Renewable Energy Consumption
  geom_line(data = consumptiondata, aes(x = Month, y = Total_Renewable_Energy_Consumption, color = "Consumption")) + 
  # Customize the plot
  labs(title = "Total Renewable Energy Production and Consumption", x = "Date", y = "Value") + 
  scale_color_manual(values = c("Production" = "blue", "Consumption" = "red"), name = "Type") +
  theme_minimal()

## Total Primary Energy

primary <- ggplot() + 
  # Plot Total Primary Energy Production
  geom_line(data = data, aes(x = Month, y = Total_Primary_Energy_Production, color = "Production")) + 
  # Plot Total Primary Energy Consumption
  geom_line(data = consumptiondata, aes(x = Month, y = Total_Primary_Energy_Consumption, color = "Consumption")) + 
  # Customize the plot
  labs(title = "Total Primary Energy Production and Consumption", x = "Date", y = "Value") + 
  scale_color_manual(values = c("Production" = "blue", "Consumption" = "red"), name = "Type") +
  theme_minimal()

# Combine all the plots
final_plot <- coal + fossil_fuels + nuclear_electric +
              hydroelectric + solar + biomass +
              renewable + primary +
              plot_layout(guides = 'collect') & 
              theme(legend.position = "bottom")

final_plot
```



On pense montrer qqch comme ça, on ajoute temp, on voit que conso plus eélece, donc importe, si trop de prod, tu exportes
on doit acheter (trop cher elec)
exporter







## Plot Data Production

```{r}
# Coal Production
ggplot(data, aes(x = Month, y = Coal_Production)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Natural Gas Dry Production
ggplot(data, aes(x = Month, y = Natural_Gas_Dry_Production)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Crude Oil Production
ggplot(data, aes(x = Month, y = Crude_Oil_Production)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Natural Gas Plant Liquids Production
ggplot(data, aes(x = Month, y = Natural_Gas_Plant_Liquids_Production)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Total Fossil Fuels Production
ggplot(data, aes(x = Month, y = Total_Fossil_Fuels_Production)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Nuclear Electric Power Production
ggplot(data, aes(x = Month, y = Nuclear_Electric_Power_Production)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Hydroelectric Power Production
ggplot(data, aes(x = Month, y = Hydroelectric_Power_Production)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Geothermal Energy Production
ggplot(data, aes(x = Month, y = Geothermal_Energy_Production)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Solar Energy Production
ggplot(data, aes(x = Month, y = Solar_Energy_Production)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Wind Energy Production
ggplot(data, aes(x = Month, y = Wind_Energy_Production)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Biomass Energy Production
ggplot(data, aes(x = Month, y = Biomass_Energy_Production)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Total Renewable Energy Production
ggplot(data, aes(x = Month, y = Total_Renewable_Energy_Production)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Total Primary Energy Production
ggplot(data, aes(x = Month, y = Total_Primary_Energy_Production)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")
```

## Plot Data Consumption

```{r}
# Coal Consumption
ggplot(consumptiondata, aes(x = Month, y = Coal_Consumption)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Natural Gas Consumption
ggplot(consumptiondata, aes(x = Month, y = Natural_Gas_Consumption)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Petroleum Consumption
ggplot(consumptiondata, aes(x = Month, y = Petroleum_Consumption)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Total Fossil Fuels Consumption
ggplot(consumptiondata, aes(x = Month, y = Total_Fossil_Fuels_Consumption)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Nuclear Electric Power Consumption
ggplot(consumptiondata, aes(x = Month, y = Nuclear_Electric_Power_Consumption)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Hydroelectric Power Consumption
ggplot(consumptiondata, aes(x = Month, y = Hydroelectric_Power_Consumption)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Geothermal Energy Consumption
ggplot(consumptiondata, aes(x = Month, y = Geothermal_Energy_Consumption)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Solar Energy Consumption
ggplot(consumptiondata, aes(x = Month, y = Solar_Energy_Consumption)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Wind Energy Consumption
ggplot(consumptiondata, aes(x = Month, y = Wind_Energy_Consumption)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Biomass Energy Consumption
ggplot(consumptiondata, aes(x = Month, y = Biomass_Energy_Consumption)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Total Renewable Energy Consumption
ggplot(consumptiondata, aes(x = Month, y = Total_Renewable_Energy_Consumption)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")

# Total Primary Energy Consumption
ggplot(consumptiondata, aes(x = Month, y = Total_Primary_Energy_Consumption)) + 
  geom_line() + 
  labs(title = "Time Series Plot", x = "Date", y = "Value")
```

## Plot Temperature

```{r}
# Temperature over the years
ggplot(temperature, aes(x = Date, y = Temperature)) + 
  geom_line() + 
  labs(title = "Temperature Time Series", x = "Date", y = "Temperature")
```

## All plots together (Production)

```{r}
ggplot(data, aes(x = Month)) + 
  geom_line(aes(y = Coal_Production, color = "Coal Production")) +
  geom_line(aes(y = Natural_Gas_Dry_Production, color = "Natural Gas Dry Production")) +
  geom_line(aes(y = Crude_Oil_Production, color = "Crude Oil Production")) +
  geom_line(aes(y = Natural_Gas_Plant_Liquids_Production, color = "Natural Gas Plant Liquids Production")) +
  geom_line(aes(y = Total_Fossil_Fuels_Production, color = "Total Fossil Fuels Production")) +
  geom_line(aes(y = Nuclear_Electric_Power_Production, color = "Nuclear Electric Power Production")) +
  geom_line(aes(y = Hydroelectric_Power_Production, color = "Hydroelectric Power Production")) +
  geom_line(aes(y = Geothermal_Energy_Production, color = "Geothermal Energy Production")) +
  geom_line(aes(y = Solar_Energy_Production, color = "Solar Energy Production")) +
  geom_line(aes(y = Wind_Energy_Production, color = "Wind Energy Production")) +
  geom_line(aes(y = Biomass_Energy_Production, color = "Biomass Energy Production")) +
  geom_line(aes(y = Total_Renewable_Energy_Production, color = "Total Renewable Energy Production")) +
  geom_line(aes(y = Total_Primary_Energy_Production, color = "Total Primary Energy Production")) +
  labs(title = "Energy Productions Time Series", x = "Date", y = "Production Value") +
  scale_color_manual(values = c("Coal Production" = "black", "Natural Gas Dry Production" = "red", 
                                "Crude Oil Production" = "blue", "Natural Gas Plant Liquids Production" = "green", 
                                "Total Fossil Fuels Production" = "orange", "Nuclear Electric Power Production" = "purple",
                                "Hydroelectric Power Production" = "brown", "Geothermal Energy Production" = "pink",
                                "Solar Energy Production" = "yellow", "Wind Energy Production" = "grey",
                                "Biomass Energy Production" = "cyan", "Total Renewable Energy Production" = "magenta",
                                "Total Primary Energy Production" = "beige")) +
  theme_minimal() +
  theme(legend.title = element_blank(), legend.position = "bottom")
```

## All plots together (Consumption)

```{r}
ggplot(consumptiondata, aes(x = Month)) + 
  geom_line(aes(y = Coal_Consumption, color = "Coal Consumption")) +
  geom_line(aes(y = Natural_Gas_Consumption, color = "Natural Gas Consumption")) +
  geom_line(aes(y = Petroleum_Consumption, color = "Petroleum Consumption")) +
  geom_line(aes(y = Total_Fossil_Fuels_Consumption, color = "Total Fossil Fuels Consumption")) +
  geom_line(aes(y = Nuclear_Electric_Power_Consumption, color = "Nuclear Electric Power Consumption")) +
  geom_line(aes(y = Hydroelectric_Power_Consumption, color = "Hydroelectric Power Consumption")) +
  geom_line(aes(y = Geothermal_Energy_Consumption, color = "Geothermal Energy Consumption")) +
  geom_line(aes(y = Solar_Energy_Consumption, color = "Solar Energy Consumption")) +
  geom_line(aes(y = Wind_Energy_Consumption, color = "Wind Energy Consumption")) +
  geom_line(aes(y = Biomass_Energy_Consumption, color = "Biomass Energy Consumption")) +
  geom_line(aes(y = Total_Renewable_Energy_Consumption, color = "Total Renewable Energy Consumption")) +
  geom_line(aes(y = Total_Primary_Energy_Consumption, color = "Total Primary Energy Consumption")) +
  labs(title = "Energy Consumptions Time Series", x = "Date", y = "Consumption Value") +
  scale_color_manual(values = c("Coal Consumption" = "black", "Natural Gas Consumption" = "red", 
                                "Petroleum Consumption" = "blue", "Total Fossil Fuels Consumption" = "green", 
                                "Nuclear Electric Power Consumption" = "purple",
                                "Hydroelectric Power Consumption" = "brown", "Geothermal Energy Consumption" = "pink",
                                "Solar Energy Consumption" = "yellow", "Wind Energy Consumption" = "grey",
                                "Biomass Energy Consumption" = "cyan", "Total Renewable Energy Consumption" = "magenta",
                                "Total Primary Energy Consumption" = "beige")) +
  theme_minimal() +
  theme(legend.title = element_blank(), legend.position = "bottom")
```

## Forecast - Coal Consumption

```{r}
# Tsibble
coal_consumption_monthly.ts <- consumptiondata %>% 
  mutate(Year_Month = yearmonth(Month)) %>%
  as_tsibble(index = Year_Month) %>%
  select(Year_Month, Coal_Consumption)

# Gaps
coal_consumption_monthly.ts <- has_gaps(coal_consumption_monthly.ts) #no gaps

# STL Decomposition
coal_consumption_monthly.ts %>%
  model(classical_decomposition(Coal_Consumption, type = "multiplicative")) %>%
  components() %>%
  autoplot() + xlab("Year")
```

Comparing the STL decomposition of the Coal Production & The Coal Consumption, we clearly observe that the overall curves are similar, both presenting similar trends and seasonality. There seems to be a higher variation in the randomness beginning in the 2010s, but more visible in the 2020s, notably due to Covid-19. This events shows that Covid impacted the American's energy consumption, mainly as they were at home.
